\chapter{Conclusion}

The contribution in this thesis was to design an exergame for elderly compatible with the motion sensor technology provided by Microsoft Kinect. This have been done by studying and analysing related theory and literature, which includes studying elderly and their attitude towards exercising, research done on elderly and exergames, and theory about video game design, usability and system design in general. We have followed one iteration of the cycle of user-centered design, all from gathering information about the users' needs and expectations, to specifying requirements and designing solutions, which we evaluate together with the users. This have been done by conducting two workshops with a group of elderly from the organisation "Seniornett". The motivation for this thesis is built upon the work presented in our project assignment \cite{project}. However, in that assignment we focused on the exergame's potential in the market, with physiotherapists as customers. This thesis will focus on the exergame's context of use, seen from the end users. 

Results from the first workshop clearly showed that existing commercial games are not developed with elderly as the intended user group. These games were at times too fast, they had to complex menus, and the games contained disturbing elements and features. In addition, there were lack of instructions and feedback, which made the informants feel that they did not master the challenges. This also made it difficult to know if what they did was done correctly. However, the commercial games held features that captivated the elderly. Well-known activities in a real-life environment enhanced the gaming experienced, and they seemed to enjoy playing.     

The exergame concept presented in this thesis is based on the use of challenges, activities and environments that are familiar to the elderly users. The story for the game is a walk in the nature, with additional, well-known activities like picking apples, swimming, paddling and wood chopping. Exercises that are shown to be good for elderly are built into the story of the game, where the game will take the player through a full body workout. The inclusion of clear instructions is emphasised, and motivational factors, like the possibility to achieve goals, are included. The exergame contains challenges to also exercise cognitive skills, like quizzes to answer along the walk in the nature. To enhance the gaming experience, the user interface is simple and not too extensive.

The evaluation of the exergame concept in the second workshop showed that the elderly liked the idea of the familiar forest environment. What the elderly were concerned about was the inclusion of quizzes while walking or performing other challenges, as they did not like the idea of doing more than one task at the same time. They were also confused about how progression with goals and levels were presented. About the menu in the game, they had no negative comments. The general evaluation of the exergame was that the elderly liked the concept presented for them. The results from the evaluation have been listed as recommendations for future work. 

The results from the workshops presented in this thesis are based upon feedback from the chosen group of informants. This is a group of fit elderly, which exercise regularly. They all have technology experience, but not with video games. Knowing elderly as a group, we evaluate the informants to be a homogeneous group based on their interests, and physical and psychological challenges. The choice of representative users is important in the work of gathering reliable information, and we have acknowledged that the inclusion of a homogeneous group might not have provided us with information representative for elderly as a group.       

Use of exergames for exercise and rehabilitation is a popular topic, and huge amount of research have been done. To make an exergame for the elderly users, it is crucial to take characteristics of elderly, as diversity in needs and interests, and the wide range of various physical and psychological challenges, into consideration. We believe the exergame concept presented in this thesis, together with the recommendations provided for future work, is a good foundation for a future exergame for elderly. The provided guidelines and requirements can help developers create user-friendly video games that meet the needs and characteristics of elderly.

[GAMMEL tekst]
During our study we found commercial exergames to be not as inappropriate for elderly as previous research first indicated. There were functionality and elements that did not suited the characteristics of elderly, but mostly, we evaluate the commercial exergames to meet some needs for effectiveness, efficiency and satisfaction.  What has to be taken into consideration when developing an exergame for this user group are real-life experiences, more instructions, motivational factors, clear goals, simple menus, technical precision, and social aspects. This thesis presents an exergame concept, that are based on a real-life theme with familiar activities and challenges, that has included most the considerations mentioned above. We believe it is a good foundation for a future exergame for elderly. In addition, our thesis provides general guidelines and requirements that can help developers create user-friendly video games that meet the needs and characteristics of elderly.  

\section{Future Work}

Future work for an exergame for elderly can head in different directions. 

One direction would be to continue working on the video game concept presented in this thesis. This will imply inclusion and potential improvement of the aspects for future work listed in Section \ref{sec:summarydiscW2}. This should be taken into a new iteration of the cycle of user centered design. Next time design is to be prototyped, there should be focus on developing more interactive prototypes to increase the users understanding of the game. 

Another direction to go would be to look into, and study and validate, our summarised guidelines, findings, and specified requirements. We feel that our study has been thorough, but we acknowledge that there always will be further research. One approach would be to conduct new workshops, with a wider selection of participants. This implies recruiting a higher number of participants than we did, but also including elderly with more diversity due to technology experience, interests, education, physical health, and life-style. A more diverse group of elderly will provide the study with more reliable information.   

A place for this exergame to be used is, as concluded in \cite{project}, at physiotherapy clinics as an assistant tool in their work. Future work should therefore be to look at this exergame from the physiotherapists point of view, to understand and specify their wants and needs. This due to both wanted functionality and requirements to an user interface. Future work should also include looking into other possibilities for how and where the exergame could be used.   

Since the technical aspects related to the delay was such an issue, it should be the future work for developers to figure out how to minimise or eliminate this. This could be grounded in choice of software. 
