\chapter{Workshops}
This chapter is about the two workshops we have conducted with participants from "Seniornett". We will present and describe the purpose of the workshops and how the workshops were set up and executed.

\section{Recruitment of Participants for the Workshop}
"Seniornett" is an organization in Norway which work to include the older generation (55 years +) in the emerging information technology and to help them gain digital competence. The organization was started in 1997 and is represented in every county in Norway \cite{seniornett}. Even though this group includes seniors who are interested in learning about new technology, something not every senior is, we decided that we wanted to start with this group for our workshop. This group may still have the same physical limitations, but may have a higher understanding of technology than other seniors. Every semester the local clubs arrange 3-4 meetings with different topics related to technology.

We contacted the manager of the club in Trondheim. He invited us to present the topic of this thesis in their next meeting, which was 02.25.13. This presentation would both serve as a contribution to this club, as well as promotion of our project. The latter was mostly to gain interest in participating in the planned workshop. The presentation slides can be found in appendix B. We will now shortly present what we presented in the meeting as well as some of the comments and questions that arose during the meeting. 

The main focus of the presentation was not our master thesis, but rather about the evolution of computer and video games, and how these can be used for different purposes. We had a brief description of our master thesis in the beginning of the presentation. After this, we introduced computer and video games. With the assumption that the audience had little knowledge or experience with this kind of technology, we started out very easy, with the description of computer games, with examples like solitaire, scrabble and chess. We also presented "The Sims" as an example of a single-player game with very innocent actions, and "Counterstrike" as an examples of a more violent game, where the player is playing over the internet alone or in a team against an enemy. Later we discussed violent computer and video games. Even though most people think of violent games as something that affect the players in a negative way, we wanted to describe the positive aspects of violent games. We used World of Warcraft as an example. Violent games often requires some kind of cooperation in a team, where each player has to understand the other team members' strengths and weaknesses and coordinate these within the team. These are skills that are important also in real-life, and people playing these type of games, has shown to have good ability to obtain the same skills in real life. We decided that a short description of this topic with a positive view, was suitable in this setting, as most focus around this topic is negative. However, we did not get any comments or feedback on this topic. 

From computer games we moved over to video games. We described how video games has emerged from being sedentary games to be more active games controlled by body movements. Because of the popularity around the SuperMario series, we decided to present two different versions of this game. We showed an old version in two-dimensional space and a newer version in three-dimensional space. This was to show the audience how games with the same concept, can change throughout the years. The audience seemed to like the video clips of SuperMario and laughed during the showing. One women asked if the clip was already recorded and where the player was in the game? We described that this was a clip that someone had recorded and published on YouTube and that the player was a human being that controlled Mario with a controller. We held up the controller for her to see. She immediately understood. It struck us that these kind of things that we, who are familiar with this kind of games, take for granted, is maybe not so intuitive for the older generation. Therefore, the next time we are talking to an audience about unfamiliar technology, we should be more precise in our descriptions. Further we showed a video clip of a car race game, called "Need for Speed". The audience also seemed to like this, but did not comment it. 

We also decided to talk about the topic "Serious Games" because exergames fall under this category. We believe a description of this gave exergames more credibility among the audience. We started by describing how games could be used in education, and we used a Norwegian mobile application, called "QuizBattle", as an example. We talked about the importance of focusing on the game story and entertainment, while keeping the pedagogics and educational factors as underlying factors. Further we described how games can be used as simulators in for example surgery, military actions and flight controlling. Finally we discussed exergames. We briefly talked about how dance pads have been taken in use in physical education at different schools. We also described how so called exergames emerged from being "just" fun and entertaining games to be more "serious" with focus on physical exercise after realising that these type of games actually gave health improvements.

From this, we naturally proceeded to describing Xbox Kinect. We described how this technology has gained acceptance within the health community, because of its accuracy and easiness, and that this is why this technology will be used in our project. We also presented the advantages and the limitations with the use of a  game like this for the elderly user, and explained why there was a need for making a customized game. One man told us that his wife, who is 72 years old, already was exercising with Nintendo Wii, and that it seemed like she enjoyed it.

To make it easier for the audience to understand what we presented, we provided a demonstration where we played Kinect Sports 2, showed on a big canvas. We played tennis and skiing. With tennis, we showed them how only one player could physically play against a computer controlled player. With skiing we showed how two players could play and compete against each other at the same time. The audience seemed to enjoy the demonstration and they seemed impressed. However, they had some concerns. One women expressed a concern when it comes to the setup of the game, and she wondered if she would need a projector. We explained that the only thing needed was the Xbox, the Kinect sensor, the games and either a television or a computer. We also admitted that the setup and introduction of the game are not very intuitive today, and that this will be something we will consider when developing a game concept. A man asked how much the game costs, and we answered that a package with the Xbox, a Kinect sensor and two games (Kinect Sports 2 and Kinect Adventures), costs about 2000 NOK. A woman expresses a concern when it comes to dizziness when playing the ski game. She thought it was too many details. She did not think these games would fit a person with decreased balance. We confirmed that it is not the first time this is mentioned about these kind of games. An other women had some comments about the music and the theme of the games. She felt that the games were too "childish" to appeal to her. She mentioned dance as a possible theme for a game, and a Norwegian folk dance called "Rørospols" where specifically mentioned. In addition, she expressed a wish for music that appealed to her generation. A third women said that she thought this game would fit best in a group setting and not necessarily at home. She expressed concerns about the space in her living room. Others from the audience agreed that the game would fit well as a group activity, because it would be more fun in that way.

After the discussion, we presented the planned workshop and invited the audience to help us with this. We told them that to make a user-friendly game, it is very important to involve the user, and that their help would mean a lot for the development of a game concept. In addition, we explained what was expected from them during the workshop and we explained that we would video record the workshop and sound record the interviews. We also made them aware that participating is voluntary and that the project is legally reported to \ac{nsd}. Everyone in the audience got a document containing information about the project and informed consent, see appendix Y. Four people signed up at that time. X people signed up later by e-mail.

After the meeting was ended we had some casual conversations with the audience. The same lady who suggested that there should be more suitable music for their generation, gave us an example from SATS, a fitness center in Norway, where she is regularly exercising. She told that they arrange weekly senior exercise groups, called "Senior Pulse", where they only used music more related to this generation's interests. An other woman expressed big concern when it comes to the pace of the game. She meant that this would be one of the biggest challenges. She suggested that a game for their generation should have the possibility to choose pace, so that the player could start slowly and speed up the pace after getting familiar with the game.

The meeting was was very useful for us. We got participants for our workshop and we got many useful questions and comment, that will be taken into account in our further work. See appendix B for more detailed description of the presentation and appendix C for more detailed notes taken during the meeting (both appendix B and C are only provided in Norwegian).
    
\section{Workshop 1}
The primary goal for our first workshop was to introduce the participants, from now on called informants, to the Xbox Kinect technology and to three commercial games. We wanted to observe how the interacted with the technology and how well they enjoyed playing, and we wanted to have a discussion where the informants could talk about how they experienced the gaming session. We also used a short survey to get to know the informants technology experience and attitude towards exercising.    

The workshop was held over a two day period, the 13th and 14th of March, with location at "Gulhuset, Voll gård", a place familiar to the informants from "Seniornett". The workshop started around 2 pm and lasted approximately three hours. We had recruited eight informants from "Seniornett" to our workshop, three males and five females. They were divided into two groups, one for each day. One of the recruited females had an accident, which made it difficult for her to participate in the workshop. We therefore ended up with seven informants. The informants average age was 70.6 years (with a standard deviation of 7,9 years). In addition to us and the informants, we had two Ph.D. students with us the first day who acted in the role as facilitators. The second day our supervisor, Lill Kristiansen, joined us and took the role as the facilitator.   

\subsection{Agenda}
In advance we had sat up an agenda for how we wanted to carry out the two workshop days, see Table \ref{tab:agenda}.  

\begin{table} [ht!]
\centering
    \begin{tabular}{|l|l|}
       \hline
       \textbf{Introduction} & 15 minutes  \\ \hline
       \textbf{Survey} & 10 minutes  \\ \hline
       \textbf{Single play} & 15 minutes for each participant \\ \hline
       \textbf{Multi play} & 10 minutes for each participant \\ \hline
	   \textbf{Group discussion} & 65 minutes \\ \hline
    \end{tabular}
    \caption[Workshop Agenda]{Workshop agenda}
    \label{tab:agenda}
\end{table} 

\subsection{Location and Equipment}
We used "Gulhuset, Voll gård" as location for our workshop. This is a location well known for the informants. This location is used for various events, like theme lectures, song meetings, and story telling gatherings. This location has coaches, tables and chairs, and it has a small kitchen where it is possible to make coffee and something to eat. "Gulhuset" was ideal to use for our workshop, not only because is is familiar to the informants, but because it was possible to make a living room-like atmosphere. It was also ideal because it is a place where we imagine that elderly could meet and play a future exercise game. At "Gulhuset" also possess a screen and a projector, which we used for our introduction. However, during the gaming session we chose to use a 42" flat screen. This was to make the gaming experience as natural as possible, as most people do not have screens and projectors in their own homes. We borrowed the flat screen from our department at NTNU. In addition to the the flat screen we had a Xbox, a Kinect sensor and three commercial games, which where use for the gaming session. This equipment was bought with support from a foundation for master thesis with computer science as subject [ENDRE litt på denne].   

In the workshop we used both video and audio recording to be able to interact with the informants. We rented a dictaphone, two video cameras (a Sony Handycam and a Panasonic 3MOS) and a rack for each camera. We wanted to use two video cameras to be able to make recordings from different angles. We had one video camera in the front of the room to capture movement and facial expressions, and we had one in the back to see movements from behind, in addition to recording interaction with the Kinect sensor and the games. The dictaphone was used for the group discussion.      

\subsection{Execution}
In the introduction we shortly presented ourselves and the background and main goal for our master thesis. We also informed about the purpose of the workshop, and presented the agenda for the day. The main part of the introduction was a review of the consent form, see Appendix X, where we highlighted important aspects as video and audio recording and that participation is voluntary. After the introduction the informants had some time to look over the consent form before they signed two copies, one for themselves and one for us. We then handed out a short survey to the informants, where they were asked a few questions about themselves, their technology experience and how they attitude are towards exercising. 

After finishing this session we were ready to start gaming. The gaming session was divided into two parts, one part where the informants played some pre-chosen games individually, and one where they played together in pairs. The order of which part that where played first was for randomisation changed from day one to day two. In workshop day one there were only three informants, so they took turns playing during the multi player part. 

We presented the informants for three different pre-chosen commercial games, \emph{Fruit Ninja}, \emph{Your Shape Fitness Evolved 2012}, and \emph{Kinect Sports Season Two}. \emph{Fruit Ninja} is a game where you have to use your arms like a ninja to slice fruit that is thrown up into the air. The goal is to slice as much delicious fruit as possible in a short period of time, without hitting any bombs that are thrown up together with the fruit. This games features simultaneous multi play with both competition and cooperation. The choice for using \emph{Fruit Ninja} in the workshop was to present the informants for a game based on pure fun and movement, with a concept far from something one might experience in real life. \emph{Your Shape Fitness Evolved 2012} is a popular exercise game for Kinect. This game has over 90 hours of activities, which can be used to design your own workout program. \emph{Your Shape Fitness Evolved 2012} follows your shape, fitness level and goals, and uses this to schedule the difficulty for the activities. You have the possibility to choose exercises for specific muscle groups, you can join classes like jump rope, cardio boxing and yoga, or you can take a virtual run in New York or Paris \cite{yourshape}. We chose this game to let the informants experience a game that are designed with exercise as the main purpose, with the use of realistic exercises. We presented a workout program for the informants that are designed specific for elderly. This program consist of a set of aerobic-like movements. \emph{Kinect Sports Season Two} is game that consist of a bundle of six sports, which are tennis, darts, base ball, American football, skiing and golf. These sports stimulates movement and activity in a fun and motivating way, though exercise is not the main focus. \emph{Kinect Sports Season Two}, depending on the sport chosen, offers both competitive and cooperative multi play. We wanted to present this game for the elderly because of its amusing and real-life activities. \emph{Your Shape Fitness Evolved 2012} and tennis from \emph{Kinect Sports Season Two} was used for single play, while \emph{Fruit Ninja} and skiing from \emph{Kinect Sports Season Two} was used for multi play. 

Before the informants started on their multi player session, we introduced how the Kinect sensor works and how to interact with it. We also describe the games they where about to play, and the goal of the games. In the two games where the informants played together, we started the games for them. One of the goals with the workshop was to observe how well the informants understood the technology and the games presented for them, so when they where about to play individually, we let them walk through the menu and start up the games alone. We guided them when necessary. 

When the informants had played individually and together in pairs we sat down and had a group discussion. We asked them open questions about their experience of the technology and the gaming session. We wanted to know if they liked the games or not, and if so why. It was also in out interest to find out if this technology is something the informants would use, and if they could imagine using it for exercise. Our questions acted just as a starting point for discussion, most of the time the informants talked freely with us and each other.                   

\section{Workshop 2}